---
title: "Feature Selection"
author: "Joan Mwangi"
date: "7/17/2021"
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Problem Definition

# Data Sourcing

```{r}
#previewing the first 6 rows
sales <- read.csv("C:/Users/jojos/Downloads/Supermarket_Dataset_1 - Sales Data (2).csv")
head(sales)
```

# Check the Data

```{r}
#checking the structure of the dataset
sapply(sales,class)
```

```{r}
class(sales)
```

Forecast belongs to class : data.frame

```{r}
dim(sales)
```

There are 1000 rows and 2 columns This dataset has 1000 rows and 16
columns

# Perform Data Cleaning 

### Checking for null values

```{r}
colSums(is.na(sales))
```

There are no null values within the dataset \#\#\#Checking for
duplicates

```{r}
duplicates <- sales[duplicated(sales),]
duplicates
```

There are no duplicates

# Perform Exploratory Data Analysis Univariate

```{r}
library(DataExplorer)
plot_bar(sales)
```

```{r}
plot_histogram(sales)
```

## Measures of central tendency 

```{r}
library(psych)
describe(sales)
```

# Implement the Solution

## Finding anomalies within the supermarket sales forecasting dataset

```{r}
#loading the libraries
library(caret)
library(corrplot)
library(lattice)
library(ggplot2)
```

```{r}
#calculating the correlation matrix
#identifying numerical columns
suppressWarnings(
        suppressMessages(if
                         (!require(corrplot, quietly=TRUE))
      install.packages("corrplot")))

```

```{r}
sales_num <- sales[sapply(sales, is.numeric)]
#installing corrplot
library(corrplot)
corr <- cor(sales_num)
#finding columns that are highly correlated
highcor <- findCorrelation(corr,cutoff=0.75)
names(sales[,highcor])
```

Columns identified with correlation above Gender,Tax and Customer Type

```{r}
#removing the features with a correlation of above 0.75
sales1 <- sales_num[-highcor]
head(sales1)
```

```{r}
#comparing correlogram with high correlation and  without
par(mfrow = c(1, 2))
corrplot(corr, order = "alphabet")
corrplot(cor(sales1), order = "alphabet")
```

# Insights

The highly correlated variables are: Gender, Tax and customer type
